High-Performance Data Analysis on JANUS using Apache Spark

### 2013-12-18 paper.tex
finished revising PageRank overview

For the PageRank algorithm, only the source url and the dest url is needed.
Therefore the dataset was was preprocessed to filter out only the necessary
information.  If the entire dataset is assumed to be represented by a huge
graph, then here each DBPedia instance maps to a Wikipedia article or a page
which inturn could be mapped to a vertex of the graph. The links between the
instances could then be mapped to edges of the graph.

@@

The scala implementation of this algorithm is as shown below:
\begin{lstlisting}[language=ML]
vallinks =// RDD of (url, neighbors) pairs
varranks =// RDD of (url, rank) pairs
for(i<- 1 to ITERATIONS) 
{
  valcontribs=links.join(ranks).flatMap{
  case (url, (links, rank)) => links.
  map( dest=> (dest, rank/links.size))
  }
  ranks =contribs.reduceByKey(_ + _).
  mapValues(0.15 + 0.85 * _)
}
ranks.saveAsTextFile(...) 
\end{lstlisting}

\emph{Bagel}:
The Bagel programming model was developed by spark.This is the implementation of the graph processing framework developed by google called the \emph{Pregel} system[??]. Bagel is still in the development stage where in more features are being added to the basic graph computation framework.

Any graph algorithm runs for several iterations and each iteration is called a Superstep.The graph is represented as a distributed dataset of $(K,V)$ pairs.  Here, the keys are vertex IDs, the values are vertex and associated states of the vertices.Every vertex in the graph will be associated with a user specified function.In each superstep, the algorithm runs the function associated with the vertex using the messages recieved from the previous superstep. The output is a new vertex state and a list of all outgoing messages which is then sent to all the other dependent vertices which is then used as input for the next iteration.

For the PageRank algorithm, as previously explained the vertices are pages and the edges are links.The list of messages generated at every iteration is the contribution of the PageRank for all its outgoing links.

Bagel provides APIs for vertices, edges and messages which could be extended to customize for any algorithm.This enables the user to write any compilcated code within few lines of code.

Further, bagel also provides combiner utility where in a reduction operation is applied on the outgoing message which reduces the communication overhead.The aggregator utility helps in reducing the message across all nodes for all vertices of the graph after every superstep.

\emph{Hash Partitioner}:The algorithm makes use of a hash partitioner which is a Controlled data partitioning utility that reduces communication cost.Also, it keeps the information of how data was partitioned across jobs for further reduction of communication cost.

\emph{Kyro serialization}: For the bagel program a special form of serialization technique was used, this  serializes the object sent over the network that will be cached in serialized form. The kryo serialization is much faster than the default java serialization used in the spark system as it the communication cost.

@@

We tested a Bagel implementation of PageRank on several of 5 million links, 30
million links, and 50 million links. We ran these computations on allocations
ranging from 4 to 16 nodes, incrementing by two.  As shown in Figure
\ref{fig:scalability}, for the scalability test, the larger the dataset is, the
better the scalability is because the ratio of computation to communication is
large. We had to run on a small dataset because of a technical problem with our
Spark installation.\footnote{
\url{https://groups.google.com/forum/#!topic/spark-users/-nsoF7oonOM}}

@@

\begin{figure}[H]
        \centering
        \includegraphics[width=90mm]{images/bagelsimple.png}
        \caption{Bagel VS. Simple PageRank}
        \label{fig:bagelsimple}
    \end{figure}

In order to measure the relative performance of the Bagel and simple
implementations, we ran both versions of PageRank on
number of nodes from 4 to 16 with increment 2 and a dataset with 64,000 pages and
30 million links in total with sufficient memory. In Figure
\ref{fig:bagelsimple} regardless of time, we can see that bagel version
outperforms simple version a little bit. The main reason is that for a smaller
dataset, the overhead of managing vertices, edges and messages are so huge that
the benefits of asynchronous background messages that Spark might offer is not
that obvious. We couldn't test on a big dataset because of the aforementioned
technical failure. However, it's worth noting that programming is easier with
general concepts of vertices, edges and messages to represent the graph.

@@

To sum up, Spark, with its in memory data store, performs great on iterative
algorithms and degrade linearly with insufficient memory. Using a partitioner,
the dataset is uniformly distributed across the nodes to avoidance of possible
stragglers. For the scalability test, the time increase with the number of nodes
is due to the communication overhead for the small dataset. We believe that
solving the akka error would enable the program to run on larger datasets, and
show great scalability nodes counts of 10 and under.

@@

% workers and master communication over spark protocol

>>>>

For the PageRank algorithm, only the \emph{source url} and the \emph{dest url}
is needed.  Therefore the dataset was was preprocessed to filter out only the
necessary information.  If the entire dataset is assumed to be represented by a
huge graph, then here each DBPedia instance maps to a Wikipedia article or a
page which in turn could be mapped to a vertex of the graph. The links between
the instances could then be mapped to edges of the graph.

@@

The Scala implementation of this algorithm is as shown below:
\begin{lstlisting}[language=ML]
vallinks =// RDD of (url, neighbors) pairs
varranks =// RDD of (url, rank) pairs
for(i<- 1 to ITERATIONS) 
{
  valcontribs=links.join(ranks).flatMap{
  case (url, (links, rank)) => links.
  map( dest=> (dest, rank/links.size))
  }
  ranks =contribs.reduceByKey(_ + _).
  mapValues(0.15 + 0.85 * _)
}
ranks.saveAsTextFile(...) 
\end{lstlisting}

\emph{Bagel}:
The Bagel programming model was developed by Spark. This is the implementation
of the graph processing framework developed by Google called the \emph{Pregel}
system[??]. Bagel is still in the development stage where in more features are
being added to the basic graph computation framework.

Any graph algorithm runs for several iterations and each iteration is called a
Superstep. The graph is represented as a distributed dataset of $(K,V)$ pairs.
Here, the keys are vertex IDs, the values are vertex and associated states of
the vertices. Every vertex in the graph will be associated with a user specified
function. In each superstep, the algorithm runs the function associated with the
vertex using the messages received from the previous superstep. The output is a
new vertex state and a list of all outgoing messages which is then sent to all
the other dependent vertices which is then used as input for the next
iteration.

For the PageRank algorithm, as previously explained the vertices are pages and
the edges are links. The list of messages generated at every iteration is the
contribution of the PageRank for all its outgoing links.

Bagel provides APIs for vertices, edges and messages which could be customized
for any algorithm. This enables the user to write any complicated code with
only few lines of code.

Further, Bagel also provides a combiner utility where in a reduction operation
it is applied on the outgoing message which reduces the communication overhead.
The aggregator utility helps in reducing the number of messages across all
nodes for all vertices of the graph after every superstep.

\emph{Hash Partitioner}: The algorithm makes use of a hash partitioner which is
a controlled data partitioning utility that reduces communication cost. Also,
it keeps the information of how data was partitioned across jobs for further
reduction of communication cost.

\emph{Kyro serialization}: For the Bagel program a special form of
serialization technique was used, this serializes the object sent over the
network that will be cached in serialized form. The Kryo serialization is much
faster than the default Java serialization used in the Spark system as it adds
to the communication cost.

@@

We tested a Bagel implementation of PageRank on several of 5 million links, 30
million links, and 50 million links. We ran these computations on allocations
ranging from 4 to 16 nodes, incrementing by two.  As shown in Figure
\ref{fig:scalability}, for the scalability test, the larger the dataset is, the
better the scalability is because the ratio of computation to communication is
large. We had to run on a small dataset because of a technical problem with our
Spark installation.\footnote{
\url{https://groups.google.com/forum/#!topic/Spark-users/-nsoF7oonOM}}

@@

\begin{figure}[H]
        \centering
        \includegraphics[width=90mm]{images/Bagelsimple.png}
        \caption{Bagel VS. Simple PageRank}
        \label{fig:Bagelsimple}
    \end{figure}

In order to measure the relative performance of the Bagel and simple
implementations, we ran both versions of PageRank on
number of nodes from 4 to 16 with increment 2 and a dataset with 64,000 pages and
30 million links in total with sufficient memory. In Figure
\ref{fig:Bagelsimple} regardless of time, we can see that Bagel version
outperforms simple version a little bit. The main reason is that for a smaller
dataset, the overhead of managing vertices, edges and messages are so huge that
the benefits of asynchronous background messages that Spark might offer is not
that obvious. We couldn't test on a big dataset because of the aforementioned
technical failure. However, it's worth noting that programming is easier with
general concepts of vertices, edges and messages to represent the graph.

@@

To sum up, Spark, with its in memory data store, performs great on iterative
algorithms and degrade linearly with insufficient memory. Using a partitioner,
the dataset is uniformly distributed across the nodes to avoid the possible
stragglers. For the scalability test, the time increase with the number of nodes
is due to the communication overhead for the small dataset. We believe that
solving the akka error would enable the program to run on larger datasets, and
show great scalability over nodes counts of 10 and under.

@@

% workers and master communication over Spark protocol
